---
title: "Data Visualization - Napoleon's March in 1812"
author: "Shakir Ahmed"
date: 2023-09-13 (YYYY-MM-DD)
---

## Setup

Initialization of packages, and downloading additional packages needed for this homework assignment.

Theme set has also been set to the perfect standard.

```{r include=FALSE}
#Installing required packages for the homework
install.packages("pacman")
install.packages("tidyverse")
install.packages("dplyr")
install.packages("ggplot2")
pacman::p_load("tidyverse")


# set theme for ggplot2
ggplot2::theme_set(ggplot2::theme_minimal(base_size = 14))

# set width of code output
options(width = 65)

# set figure parameters for knitr
knitr::opts_chunk$set(
  fig.width = 7,        # 7" width
  fig.asp = 0.618,      # the golden ratio
  fig.retina = 3,       # dpi multiplier for displaying HTML output on retina
  fig.align = "center", # center align figures
  dpi = 300             # higher dpi, sharper image
)

```

## Napoleon's march in 1812

This exercise i to recreate the Napoleon's march plot by Charles John Minard in ggplot2. The data is provided in package 'HistData', and this is the package from where we will be importing the data.

-   Importing/reading the code using the rds() function after setting file path to 'data' folder

```{r import,warning = FALSE, message=FALSE,include=FALSE}

    library(ggplot2)
    library(scales)     # additional formatting for scales
    library(grid)       # combining plots
    library(gridExtra)  # combining plots
    library(dplyr)      # tidy data manipulations
    install.packages("HistData")
```

Now, we can recreate the plot of Napoleon's march in 1812. Parameters included are Latitude, Longitude, troop strength (size) and temperature.

```{r Napoleon_march,warning = FALSE, message=FALSE}

  
    #Troop data
    data(Minard.troops, package="HistData")
    str(Minard.troops)
    #Cities Data
    data(Minard.cities, package="HistData")
    str(Minard.cities)
    #Temperature data
    data(Minard.temp, package="HistData")
    str(Minard.temp)

    #Below plot is for plotting the latitude and longitude the troops travelled
    #Width is set as the number of survivors to show the strength of the army
    #Knitr function is used to scale the graph
     breaks <- c(1, 2, 3) * 10^5 #New vector setis defined
    ggplot(Minard.troops, aes(long, lat)) +
            geom_path(aes(size = survivors, colour = direction, group = group),
                      lineend="round") +
        scale_size("Survivors", range = c(1,10), #c(0.5, 15),
                   breaks=breaks, labels=scales::comma(breaks)) +
        scale_color_manual("Direction", 
                           values = c("cornsilk3", "black"), 
                           labels=c("Advance", "Retreat")) 

    #ggplot trick to assign it to a variable to use later
    plot_troops <- last_plot()


    #Cities are now appended on top of the above plot
    plot_troops + geom_text(data = Minard.cities, aes(label = city), size = 3)

    #Package ggrepel is used to automatically move the labels away from points and to ensure none of the labels overlap
    if (!require(ggrepel)) {install.packages("ggrepel"); require(ggrepel)}
    library(ggrepel)
    plot_troops +   
        geom_point(data = Minard.cities) +
        geom_text_repel(data = Minard.cities, aes(label = city))

    #ggplot trick to assign it to a variable to use later
    plot_troops_cities <- last_plot()

    #Plotting the temperature with custom size
    ggplot(Minard.temp, aes(long, temp)) +
        geom_path(color="grey", size=1.5) +
        geom_point(size=2)

    #Below code is to standardize the temperature format by combining temperature and date
    Minard.temp <- Minard.temp %>%
        mutate(label = paste0(temp, "° ", date))
    head(Minard.temp$label)

    #These plots are pushing the right most two points outside the graph. Using geom_text_repel to bring back the data
    ggplot(Minard.temp, aes(long, temp)) +
        geom_path(color="grey", size=1.5) +
        geom_point(size=1) +
      geom_text_repel(aes(label=label), size=2.5)

    #ggplot trick to assign it to a variable to use later
    plot_temp <- last_plot()


    #Now, we have to assemble the two graphs one below the other
    grid.arrange(plot_troops_cities, plot_temp)

    #In troops and cities plot, we have to set X axis limits to a range that will coincide with those for the plot of temperature
    #Remove X and Y axis labels scales
    #Remove legends for survivors and directions
    plot_troops_cities +
      coord_cartesian(xlim = c(24, 38)) +
      labs(x = "Longitude", y = "Latitude",title="Minard's version of Napoleon's March") +
      guides(color = FALSE, size = FALSE)

    #ggplot trick to assign it to a variable to use later
    plot_troops_cities_fixed <- last_plot() 

    #Setting theme for temperature graph
    plot_temp + 
      coord_cartesian(xlim = c(24, 38)) +
      labs(x="Napoleon's March Duration", y="Temperature",caption = "Michael Friendly (2018), Minard meets ggplot2") + 
      theme_bw() +
      theme(panel.grid.major.x = element_blank(),
            panel.grid.minor.x = element_blank(),
            panel.grid.minor.y = element_blank(),
            axis.text.x = element_blank(), axis.ticks = element_blank(),
            panel.border = element_blank())

    #ggplot trick to assign it to a variable to use later
    plot_temp_fixed <- last_plot()

    #Function to combine both the plots into one single plot and add a border
    grid.arrange(plot_troops_cities_fixed, plot_temp_fixed, nrow=2, heights=c(3.5, 1.2))
    grid.rect(width = .99, height = .99, gp = gpar(lwd = 2, col = "gray", fill = NA))
```
